---
title: "Using `theme_gla()`"
author: "Libby Rogers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the ggla theme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 6,
  fig.height = 4,
  comment = ""
)
```

This vignette summarises how to use the custom theme included in `gglaplot`.

Themes control the overall look of the non-data components of plots. `ggplot2` includes various default themes, as well as a framework for editing and building your own, on which `theme_gla` is built.

##How to apply  `theme_gla` to a `ggplot` object


```{r}
library(ggplot2)
library(gglaplot)
```

To illustrate how `theme_gla` can be used with `ggplot` plots, we'll use a dataset comparing the gender pay gap in London and the rest of the UK. Here's the dataset plotted in standard `ggplot`.



```{r}
plot = ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG, group = location, 
                                   colour = location)) +
  geom_line()
plot
```


After installing `gglaplot` the custom theme can be accessed via the `theme_gla()` function.
The function has been set up in such a way that it can be used without any arguments. The theme can be added onto any standard ggplot object in the same way as the themes built into `ggplot`.


```{r}
plot + 
  theme_gla()
```

With the default settings, `theme_gla()` changes some basic aspects of the look of the graph, including legend style and location, background colours and gridlines.

Here's the full argument and default specification for `theme_gla()`:

```{r, echo = FALSE}
str(theme_gla)
```

To change some of the more complicated aspects of the style of the plot, we need to change some of the `theme_gla()` arguments from their defaults.

##Settings in the `theme_gla()` function

###Light or Dark Theme


```{r}
plot + 
  theme_gla(gla_theme = 'dark')
```

###Gridlines

`xgridlines` controls whether gridlines appear on the x-axis of the plot.

```{r}
plot +
  theme_gla(xgridlines = TRUE)
```


###Legend

`legend` controls whether the legend appears. Where possible data should be labelled directly instead.

```{r}
plot +
  theme_gla(xgridlines = TRUE, legend = FALSE)
```

###Font size

`base_size` changes the base font size of the plot. The font size of any titles and captions will scale accordingly. Try and get the `base_size` to match the font size of the document/presentation the plot is being used in.

```{r}
plot +
  theme_gla(base_size = 20)
```

##How to use `theme_gla()`

As above, you can just add `theme_gla()` to each of your ggplot objects. It's best to add it directly after the call to `ggplot` so it won't overwrite any manual changes you make later on. As with other `ggplot` themes you can make manual changes to them directly. For example, if for a particular plot you don't want any y-gridlines you can remove them using the `ggplot::theme()` function:

```{r}
ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG, group = location, 
                                   colour = location)) +
  theme_gla(xgridlines = FALSE) +
  geom_line() +
  theme(panel.grid.major.y = element_blank())
```

You can get more information on how to manually adjust themes in the [ggplot documentation](https://ggplot2.tidyverse.org/reference/theme.html). We've tried to include the suggestions of the style guide in `theme_gla()`, but there may be plots it doesn't quite work for.

Instead of adding `theme_gla()` to every one of your plots you can also set the theme at the start of your session.

```{r}
theme_set(theme_gla(xgridlines = FALSE))
          
plot
```




