---
title: "Using `theme_gla()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the ggla theme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "",
  dpi = 300
)
```

This vignette summarises how to use the custom theme included in `gglaplot`.

Themes control the overall look of the non-data components of plots. `ggplot2` includes various default themes, as well as a framework for editing and building your own, on which `theme_gla` is built.

## How to apply  `theme_gla` to a `ggplot` object


```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(gglaplot)
```

To illustrate how `theme_gla` can be used with `ggplot` plots, we'll use a dataset comparing the gender pay gap in London and the rest of the UK. Here's the dataset plotted in standard `ggplot`.



```{r}
plot <- ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG, group = location,
                                   colour = location)) +
  geom_line()
plot
```


After installing `gglaplot` the custom theme can be accessed via the `theme_gla()` function.
The function has been set up in such a way that it can be used without any arguments. The theme can be added onto any standard ggplot object in the same way as the themes built into `ggplot`.


```{r}
plot +
  theme_gla()
```

With the default settings, `theme_gla()` changes some basic aspects of the look of the graph, including legend style and location, background colours and gridlines.

Here's the full argument and default specification for `theme_gla()`:

```{r, echo = FALSE}
str(theme_gla)
```

To change some of the more complicated aspects of the style of the plot, we need to change some of the `theme_gla()` arguments from their defaults.

## Settings in the `theme_gla()` function

### Default or Inverse Theme


```{r}
plot +
  theme_gla(gla_theme = "inverse")
```

### Axis titles

If needed, axis titles can be added to plots with the `x_axis_title` and `y_axis_title` arguments. If you're including a y-axis label you may need to adjust the length of the lines under the y-axis labels

```{r}
plot +
  theme_gla(x_axis_title = TRUE, y_axis_title = TRUE, y_label_length = 25)
```

### Gridlines

`xgridlines` controls whether gridlines appear on the x-axis of the plot.

```{r}
plot +
  theme_gla(xgridlines = TRUE)
```

### Legend

`legend` controls whether the legend appears. Where possible data should be labelled directly instead.

```{r}
plot +
  theme_gla(xgridlines = TRUE, legend = FALSE)
```

If you have particularly long labels in your legend, it can be split over multiple rows by adjusting the `guides` settings in your plot:

```{r}
plot +
  theme_gla() +
  guides(col = guide_legend(nrow = 2))
```

### Font size

`base_size` changes the base font size of the plot. The font size of any titles and captions will scale accordingly. Try and get the `base_size` to match the font size of the document/presentation the plot is being used in.

```{r}
plot +
  theme_gla(base_size = 20)
```

### Faceting

Faceting creates small multiples of plots, split into rows and columns by different variables.

Simple facet plots can use `theme_gla()` with no adjustments.

```{r}
ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG)) +
  facet_grid(cols = vars(location)) +
  geom_line() +
  theme_gla()
```

For facet plots which have data in different ranges, free scales may be needed to avoid the data in some facets being squashed. If this is the case the `free_y_facets` argument in `theme_gla` should be set to `TRUE`. This changes the position and alignment of the y-axis labels to a format which works better when the y-axis labels may vary a lot in length.

```{r}
ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG)) +
  facet_grid(rows = vars(location), scales = "free_y") +
  geom_line() +
  theme_gla(free_y_facets = TRUE)
```

Each facet can have it's own set of axis using `facet_wrap`. These can either be completely *free*, (i.e. have different ranges), or the ranges can be set using limits in `scale_y_continuous`.
```{r}
ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG)) +
  facet_wrap(facets = "location", scales = "free_y") +
  geom_line() +
  theme_gla(free_y_facets = TRUE) +
  scale_y_continuous(limits = c(0, 30))
```

If there are many small plots in a facet plot, it's worth considering scaling down the `base_size` in `theme_gla` to stop text looking out of proportion, and to reduce the `size` of some geoms to avoid losing detail.

## How to use `theme_gla()`

As above, you can just add `theme_gla()` to each of your ggplot objects. It's best to add it directly after the call to `ggplot` so it won't overwrite any manual changes you make later on. As with other `ggplot` themes you can make manual changes to them directly. For example, if for a particular plot you don't want any y-gridlines you can remove them using the `ggplot::theme()` function:

```{r}
ggplot(data = LDNUK, mapping = aes(x = Year, y = GPG, group = location,
                                   colour = location)) +
  theme_gla(xgridlines = FALSE) +
  geom_line() +
  theme(panel.grid.major.y = element_blank(),
        axis.ticks.y = element_line(colour = "white"))
```

You can get more information on how to manually adjust themes in the [ggplot documentation](https://ggplot2.tidyverse.org/reference/theme.html). We've tried to include the suggestions of the style guide in `theme_gla()`, but there may be plots it doesn't quite work for.

Instead of adding `theme_gla()` to every one of your plots you can also set the theme at the start of your session.

```{r}
theme_set(theme_gla())

plot
```




