---
title: "Using plotly with gglaplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using plotly with gglaplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "",
dpi = 300,
fig.width = 9,
fig.height = 7
)
```

The R package for plotly, an interactive plotting library, includes a function to convert ggplots into interactive ploty plots. A version of this is included in gglaplot which adjust the look of the plot to be more inkeeping with the GLA IU style guide (if theme_gla() has been used on the ggplot).

The conversion of ggplots to plotlys using `ggplotly` isn't perfect and is still undergoing development. `gglaplotly` provides only basic adjustments to this, and is experimental at this stage. The main limitations are:

* It can only be used on in built ggplot2 geoms (not any of the custom gglaplot ones like `ggla_line` or `ggla_sf`)
* Not all features of a ggplot can be converted (e.g. subtitles and captions are lost in the conversion)

For more complex figures it could be easier to start with base plotly, details on how to do this can be found [here](https://plotly-r.com/) and [here](https://plotly.com/r/).


```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(gglaplot)
library(plotly)
library(sf)
library(scales)
```

Here's a simple ggplot line chart:
```{r}
pal <- gla_pal(gla_theme = "default", palette_type = "highlight", n = c(1, 1))

plot <- ggplot(data = LDNUK,
               mapping = aes(x = Year, y = GPG,
                             group = location, colour = location)) +
  scale_y_continuous(labels = dollar_format(prefix = "", suffix = "%",
                                            accuracy = 0.1)) +
  labs(title = "Gender Pay Gap - Total (Median) - London VS UK") +
  scale_colour_manual(values = pal) +
  geom_line() +
  # Use theme_gla() - a lot of the theme styling will be carried across automatically
  theme_gla()


plot
```

Feeding this plot into `gglaplotly` will produce a basic plotly plot, including a default hovertool and basic interactivity:

```{r}
gglaplotly(plot)
```

These can then be adjusted as you would any other plotly plot. See [here](https://plotly-r.com/) and [here](https://plotly.com/r/) for more details.

The title and legend can look a bit crowded (the positioning of these outside the plot so they don't overlap is tricky). In some sitations it might be better to write the plot title in the rmarkdown document instead and not include it in the plot itself.

To adjust what it shown in the hover tool add a `text` aesthetic to the ggplot plot and reference it in `tooltip` in gglaplotly.

```{r}
plot <- ggplot(data = LDNUK,
               mapping = aes(x = Year, y = GPG,
                             group = location, colour = location,
                             text = comma(x = GPG, accuracy = 0.5, suffix = "%"))) +
  scale_colour_manual(values = pal) +
  geom_line() +
  theme_gla() +
  scale_y_continuous(labels = dollar_format(prefix = "", suffix = "%",
                                            accuracy = 0.1)) +
  labs(title = "Gender Pay Gap - Total (Median) - London VS UK")
gglaplotly(plot, tooltip = "text")
```

Range sliders can be added to view certain periods of data

```{r}
gglaplotly(plot, tooltip = "text", dynamicTicks = TRUE) %>%
  layout(xaxis = list(rangeslider = list(type = "date")))

```


`gglaplotly` also includes some default settings to tidy up the rendering of maps:

```{r, fig.cap="Contains Ordnance Survey data Crown copyright and database right [2015]", fig.align="left"}
pal <- gla_pal(gla_theme = "default", palette_type = "diverging",
               main_colours = c("blue", "red"), n = 6, remove_margin = "both")
EURef <- EURef %>%
  st_as_sf()
plot <- ggplot(data = EURef) +
  geom_sf(mapping = aes(geometry = geometry, fill = CatLeave, text = CatLeave),
          color = gla_default$background) +
  scale_fill_manual(values = pal) + 
  labs(title = "Leave Vote in the 2016 EU Referendum") +
  theme_gla()

gglaplotly(plot,
           tooltip = "text")
```

**Note**: As captions cannot be carried over from ggplots, a figure caption has been added in RMarkdown (using `fig.cap=` in the chunk header). It is also possible to add an annotation to a plotly figure using [`layout(annotation = list(...))`](https://plotly.com/r/reference/#layout-annotations), but the positioning will depend on the other features of the figure.
